Multiple Dice Working as One: CAD Flows and Routing Architectures for Silicon Interposer FPGAs

### 2015-07-05 doc/journal/fpgapaperfinal/FPGAPaper/response_to_reviewers.txt

So, to produce Table I, in principle, the routability-driven router should have been chosen; however, if so, then the cost function used by the routability-driven router needs to be modified as well in order to model the interposer and interface at the cut lines.

This leads to a few interdependent questions:
(1) Which router was used?
(2) If the timing-driven router was used, is it justifiable to report W_min values? Should the routability-driven router (enhanced as needed) have been used instead?
(3) If the routability-driven router was used, how were the cost functions enhanced?

[We used the VPR timing-driven router for all results in this paper, and believe it is the correct choice as it produces a better combination of routability and timing than VPR's routability-driven router, and its algorithm is more scalable to large circuits.

  The timing-driven router in VPR optimizes for both timing and routability (timing is optimized as much as possible, but the cost of congestion grows to eventually swamp the timing terms in congested parts of the design if necessary to achieve routability).  Hence its routability is good; in "Architecture and CAD for Deep Submicron FPGAs" by Betz, Marquardt and Rose, for example, the timing driven router was found to produce circuits with a 6% higher Wmin, but that had an operating frequency 2.58x higher.  Hence it achieves a superior combination of routability and timing than the routability-driven router, although you are correct that the routability-driven router produces slightly lower absolute channel widths.  A further problem with the routability-driven router in VPR is that its algorithm does not scale as well as the timing-driven router, and hence its runtime will be much higher and likely impractical on the larger circuits we use.   On the largest circuit for which results are presented in "Architecture and CAD for Deep-Submicron FPGAs" the VPR routability-driven router requires 14.6x more CPU time than the timing-driven router, and the CPU time gap increases with circuit size.  Hence on the much larger circuits (>10x larger) we use the CPU time of the routability-driven router would make it very difficult to use.

### 2015-07-06 doc/journal/fpgapaperfinal/FPGAPaper/partitioning.tex
Added reference on hypergraph to graph transformation. Added section for cut critical connections.
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4673 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

We explored several ways to generate the per-net subgraphs, varying the graph topology and the edge weight scheme. We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}. We assigned the same edge weight to every edge of the generated subgraph. Edge weights were computed based on the number of vertices in the originating hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.  Our intuition is that assigning lower edge weight to high-fanout nets may be beneficial, and we selected the edge weights accordingly. We refer to the combination of graph topology and edge weight scheme as the \emph{hyperedge model}.

@@

>>>>

We explored several ways to generate the per-net subgraphs, varying the graph topology and the edge weight scheme. We considered two graph topologies, clique and star, as illustrated in Figure \ref{fig:star_clique}. \hl{Ref.\mbox{\cite{hypergraph_to_graph_survey}}
considers the addition of ``dummy'' nodes to model a hypergraph as a graph, with the objective of obtaining equivalent edge cut weights after partitioning. In our star model, we preferred to designate the net source as the center of the star, rather than introduce a dummy node. This choice was motivated by a desire to optimize timing -- the star topology clearly differentiates sources from sinks, and we expect this to guide the partitioner to keep sources and sinks in the same partition.}

We assigned the same edge weight to every edge of the generated subgraph. Edge weights were computed based on the number of vertices in the originating hyperedge, $n$, and we considered edge weights equal to $1$, $1/n$, and $1/n^2$.  Our intuition is that assigning lower edge weight to high-fanout nets may be beneficial, and we selected the edge weights accordingly. We refer to the combination of graph topology and edge weight scheme as the \emph{hyperedge model}.

@@

\subsection{\hl{Interposer Wire Utilization}}
\hl{Table \mbox{\ref{table:interposer_wire_utilization}} shows the number of connections on the critical path and how many of those critical connections cross the interposer boundary. It also shows the number of cut nets per vertical channel. This provides a lower bound on the number of wires required in an interposer channel. The data was obtained at 70\% wires cut.}
\begin{table}[!htbp]
\centering
\caption{Interposer wire utilization}
\begin{tabular}{|l|r|r||r|r|}
\hline
 & \multicolumn{2}{c||}{Critical Connections} & \multicolumn{2}{c|}{Cut Nets} \\
 & Cut & Total & Total & Per Channel \\ \hline
bgm & 1 & 373 & 622 & 8.89 \\ \hline
stereovision0 & 3 & 57 & 985 & 25 \\ \hline
stereovision1 & 1 & 21 & 876 & 21 \\ \hline
LU8PEEng & 0 & 1661 & 1795 & 20 \\ \hline
mkDelayWorker32B & 4 & 58 & 771 & 17 \\ \hline
mcml & 5 & 1443 & 2891 & 28 \\ \hline
LU32PEEng & 0 & 1664 & 1130 & 5 \\ \hline
\end{tabular}
\label{table:interposer_wire_utilization}
\end{table}

\hl{Though our choice of the star topology benefits timing, our partitioning flow is not timing-driven. We would expect a timing-driven partitioner to reduce the number of cut critical connections, but obtaining good critical path estimates at the partitioning stage is nontrivial\mbox{\cite{hutton}}.}
